---
title: "A Notebook for Filtering out Factors that are Less Relevant for the Analysis"
author: "Lilian Ngweta"
date: "09/27/2019"
output: html_notebook
---
Note: Re-used some of the code written by: Yuxuan Wang, Shengjin Li, and Ziyi Wang




Importing Source.R so it is possible to use variables such chr.data.2019 and cdc.data that are defined in Source.R
```{r}
source("Source.R")

```


Filtering out social determinants that are less relevant for Deaths of Despair using Benjamin Hochberg's method for multiple hypothesis testing.
```{r}
state = "US"
death_cause_despair = "Despair"
mort.cluster.raw <- function(){
  n.clusters <- 6
  cluster.counties(cdc.mort.mat(cdc.data, "US", death_cause_despair),
                   cluster.method="kmeans",
                   cluster.num=n.clusters)
}
# Weighed Avg by UNORDERED cluster
# Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  
  # Notes:
  #   - The cluster labels are UNORDERED
mort.avg.cluster.raw <- function(){
  head(cdc.data)
  cdc.data %>%
    dplyr::filter(death_cause == death_cause_despair) %>%
    dplyr::right_join(mort.cluster.raw(), by = "county_fips") %>%
    dplyr::group_by(period, cluster) %>%
    dplyr::summarise(
      death_rate = sum(death_num) / sum(population) * 10^5,
      count = n()
    ) %>% 
    dplyr::ungroup()
}


# MAPPING from UNORDERED mortality trend label to ORDERED mortality trend label
 # Variables:
  #   - ord
  # Notes:
  #   - This is a mapping from raw cluster label to ORDERED cluster.
  #       Row names are the original cluster and `ord` are the reordered cluster
mort.cluster.map <- function(){
  mort.avg.cluster.raw() %>% 
    dplyr::filter(period == "2015-2017") %>%
    dplyr::arrange(death_rate) %>% 
    dplyr::mutate(ord = as.character(1:n())) %>% 
    dplyr::select(-c(period, death_rate)) %>% 
    textshape::column_to_rownames("cluster")
}

# ORDERED mortality trend cluster label calculation
# Variables:
  #   - county_fips
  #   - cluster
mort.cluster.ord <- function(){
  order.county.clusters(mort.cluster.raw(), mort.cluster.map())
}


# Weighed Avg by ORDERED cluster
# Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  
  # Notes:
  #   - The cluster labels are ORDERED
mort.avg.cluster.ord <- function(){
  dplyr::mutate(mort.avg.cluster.raw(), cluster = mort.cluster.map()[cluster, "ord"])
}


#Extracting the mortality rates for all states and counties for years "2015-2017"
mort.rate <- function(){
  cdc.data %>% dplyr::filter(
    death_cause == death_cause_despair,
    period == "2015-2017"
  ) %>%
    dplyr::mutate(
    ) %>%
    dplyr::select(county_fips, death_rate)
  
}
# Kendall Correlation Between raw mortality rates and county health rankings social determinants
kendall.cor <- mort.rate() %>% 
      dplyr::mutate(VAR = death_rate) %>%
      kendall.func(chr.data.2019) %>%
      dplyr::mutate(
        DIR = dplyr::if_else(
          kendall_cor <= 0,
          "Protective",
          "Destructive"
        )
        #chr_code = chr.namemap.2019[chr_code, 1]
      ) %>% na.omit() %>% 
      dplyr::arrange(desc(kendall_cor)) %>% 
      dplyr::mutate(chr_code = reorder(chr_code, kendall_cor))

selected_SDs_despair <- cbind(kendall.cor, state = state)

#using Benjamin Hochberg's p.adjust() method to adjust P-values for multiple hypothesis testing to filter out factors that are less relevant.
corrected_pvalues <- p.adjust(selected_SDs_despair$kendall_p)
selected_SDs_despair <- cbind(selected_SDs_despair, Corrected_P_Vals = corrected_pvalues)
selected_SDs_despair <- selected_SDs_despair %>%
  dplyr::filter(Corrected_P_Vals < 0.05) %>% 
  dplyr::arrange(desc(kendall_cor))

#Printing out the few selected factors for deaths of despair.
print(selected_SDs_despair)
```



Filtering out social determinants that are less relevant for Cancer using Benjamin Hochberg's method for multiple hypothesis testing.
```{r}
state = "US"
death_cause_cancer = "Cancer"

# UNORDERED mortality trend cluster label calculation
# Variables:
  #   - county_fips
  #   - cluster
mort.cluster.raw <- function(){
  n.clusters <- 6
  cluster.counties(cdc.mort.mat(cdc.data, "US", death_cause_cancer),
                   cluster.method="kmeans",
                   cluster.num=n.clusters)
}


# Weighed Avg by UNORDERED cluster
# Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  
  # Notes:
  #   - The cluster labels are UNORDERED
mort.avg.cluster.raw <- function(){
  head(cdc.data)
  cdc.data %>%
    dplyr::filter(death_cause == death_cause_cancer) %>%
    dplyr::right_join(mort.cluster.raw(), by = "county_fips") %>%
    dplyr::group_by(period, cluster) %>%
    dplyr::summarise(
      death_rate = sum(death_num) / sum(population) * 10^5,
      count = n()
    ) %>% 
    dplyr::ungroup()
}


# MAPPING from UNORDERED mortality trend label to ORDERED mortality trend label
 # Variables:
  #   - ord
  
  # Notes:
  #   - This is a mapping from raw cluster label to ORDERED cluster.
  #       Row names are the original cluster and `ord` are the reordered cluster
  
mort.cluster.map <- function(){
  mort.avg.cluster.raw() %>% 
    dplyr::filter(period == "2015-2017") %>%
    dplyr::arrange(death_rate) %>% 
    dplyr::mutate(ord = as.character(1:n())) %>% 
    dplyr::select(-c(period, death_rate)) %>% 
    textshape::column_to_rownames("cluster")
}


# ORDERED mortality trend cluster label calculation
 # Variables:
  #   - county_fips
  #   - cluster
mort.cluster.ord <- function(){
  order.county.clusters(mort.cluster.raw(), mort.cluster.map())
  
}


# Weighed Avg by ORDERED cluster
# Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  
  # Notes:
  #   - The cluster labels are ORDERED
mort.avg.cluster.ord <- function(){
  dplyr::mutate(mort.avg.cluster.raw(), cluster = mort.cluster.map()[cluster, "ord"])
}


#Extracting the mortality rates for all states and counties for years "2015-2017"
mort.rate <- function(){
  
  cdc.data %>% dplyr::filter(
    death_cause == death_cause_cancer,
    period == "2015-2017"
  ) %>%
    dplyr::mutate(
    ) %>%
    dplyr::select(county_fips, death_rate)
  
}


# Kendall Correlation Between raw mortality rates and county health rankings social determinants
kendall.cor <- mort.rate() %>% 
      dplyr::mutate(VAR = death_rate) %>%
      kendall.func(chr.data.2019) %>%
      dplyr::mutate(
        DIR = dplyr::if_else(
          kendall_cor <= 0,
          "Protective",
          "Destructive"
        )
        #chr_code = chr.namemap.2019[chr_code, 1]
      ) %>% na.omit() %>% 
      dplyr::arrange(desc(kendall_cor)) %>% 
      dplyr::mutate(chr_code = reorder(chr_code, kendall_cor))

selected_SDs_cancer <- cbind(kendall.cor, state = state)

#using Benjamin Hochberg's p.adjust() method to adjust P-values for multiple hypothesis testing to filter out factors that are less relevant.
corrected_pvalues <- p.adjust(selected_SDs_cancer$kendall_p)
selected_SDs_cancer <- cbind(selected_SDs_cancer, Corrected_P_Vals = corrected_pvalues)
selected_SDs_cancer <- selected_SDs_cancer %>%
  dplyr::filter(Corrected_P_Vals < 0.05) %>% 
  dplyr::arrange(desc(kendall_cor))

#Printing out the few selected factors for cancer.
print(selected_SDs_cancer)
```


Filtering out social determinants that are less relevant for Assault using Benjamin Hochberg's method for multiple hypothesis testing.
```{r}
state = "US"
death_cause_assault = "Assault"


# UNORDERED mortality trend cluster label calculation
# Variables:
  #   - county_fips
  #   - cluster

mort.cluster.raw <- function(){
  n.clusters <- 6
  cluster.counties(cdc.mort.mat(cdc.data, "US", death_cause_assault),
                   cluster.method="kmeans",
                   cluster.num=n.clusters)
}


# Weighed Avg by UNORDERED cluster
# Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  
  # Notes:
  #   - The cluster labels are UNORDERED
mort.avg.cluster.raw <- function(){
  head(cdc.data)
  cdc.data %>%
    dplyr::filter(death_cause == death_cause_assault) %>%
    dplyr::right_join(mort.cluster.raw(), by = "county_fips") %>%
    dplyr::group_by(period, cluster) %>%
    dplyr::summarise(
      death_rate = sum(death_num) / sum(population) * 10^5,
      count = n()
    ) %>% 
    dplyr::ungroup()
}


# MAPPING from UNORDERED mortality trend label to ORDERED mortality trend label
mort.cluster.map <- function(){
  mort.avg.cluster.raw() %>% 
    dplyr::filter(period == "2015-2017") %>%
    dplyr::arrange(death_rate) %>% 
    dplyr::mutate(ord = as.character(1:n())) %>% 
    dplyr::select(-c(period, death_rate)) %>% 
    textshape::column_to_rownames("cluster")
}


# ORDERED mortality trend cluster label calculation
# Variables:
  #   - county_fips
  #   - cluster
mort.cluster.ord <- function(){
  order.county.clusters(mort.cluster.raw(), mort.cluster.map())
}



# Weighed Avg by ORDERED cluster
# Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  
  # Notes:
  #   - The cluster labels are ORDERED
mort.avg.cluster.ord <- function(){
  dplyr::mutate(mort.avg.cluster.raw(), cluster = mort.cluster.map()[cluster, "ord"])
}


#Extracting the mortality rates for all states and counties for years "2015-2017"
mort.rate <- function(){
  cdc.data %>% dplyr::filter(
    death_cause == death_cause_assault,
    period == "2015-2017"
  ) %>%
    dplyr::mutate(
      # death_rate = death_num / population * 10^5
      #death_rate = cut(death_rate, bin.geo.mort(death_cause_assault))
    ) %>%
    dplyr::select(county_fips, death_rate)
  
}

# Kendall Correlation Between raw mortality rates and county health rankings social determinants
kendall.cor <- mort.rate() %>% 
      dplyr::mutate(VAR = death_rate) %>%
      kendall.func(chr.data.2019) %>%
      dplyr::mutate(
        DIR = dplyr::if_else(
          kendall_cor <= 0,
          "Protective",
          "Destructive"
        )
        #chr_code = chr.namemap.2019[chr_code, 1]
      ) %>% na.omit() %>% 
      dplyr::arrange(desc(kendall_cor)) %>% 
      dplyr::mutate(chr_code = reorder(chr_code, kendall_cor))

selected_SDs_assault <- cbind(kendall.cor, state = state)

#using Benjamin Hochberg's p.adjust() method to adjust P-values for multiple hypothesis testing to filter out factors that are less relevant.
corrected_pvalues <- p.adjust(selected_SDs_assault$kendall_p)
selected_SDs_assault <- cbind(selected_SDs_assault, Corrected_P_Vals = corrected_pvalues)
selected_SDs_assault <- selected_SDs_assault %>%
  dplyr::filter(Corrected_P_Vals < 0.05) %>% 
  dplyr::arrange(desc(kendall_cor))

#Printing out the few selected factors for assault
print(selected_SDs_assault)
```

Filtering out social determinants that are less relevant for Cardiovascular deaths using Benjamin Hochberg's method for multiple hypothesis testing.
```{r}
state = "US"
death_cause_cardiovascular = "Cardiovascular"

# UNORDERED mortality trend cluster label calculation
# Variables:
  #   - county_fips
  #   - cluster
mort.cluster.raw <- function(){
  n.clusters <- 6
  cluster.counties(cdc.mort.mat(cdc.data, "US", death_cause_cardiovascular),
                   cluster.method="kmeans",
                   cluster.num=n.clusters)
}


# Weighed Avg by UNORDERED cluster
# Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  
  # Notes:
  #   - The cluster labels are UNORDERED
mort.avg.cluster.raw <- function(){
  head(cdc.data)
  cdc.data %>%
    dplyr::filter(death_cause == death_cause_cardiovascular) %>%
    dplyr::right_join(mort.cluster.raw(), by = "county_fips") %>%
    dplyr::group_by(period, cluster) %>%
    dplyr::summarise(
      death_rate = sum(death_num) / sum(population) * 10^5,
      count = n()
    ) %>% 
    dplyr::ungroup()
}


# MAPPING from UNORDERED mortality trend label to ORDERED mortality trend label
# Variables:
  #   - ord
  
  # Notes:
  #   - This is a mapping from raw cluster label to ORDERED cluster.
  #       Row names are the original cluster and `ord` are the reordered cluster
mort.cluster.map <- function(){
  mort.avg.cluster.raw() %>% 
    dplyr::filter(period == "2015-2017") %>%
    dplyr::arrange(death_rate) %>% 
    dplyr::mutate(ord = as.character(1:n())) %>% 
    dplyr::select(-c(period, death_rate)) %>% 
    textshape::column_to_rownames("cluster")
}


# ORDERED mortality trend cluster label calculation
 # Variables:
  #   - county_fips
  #   - cluster
mort.cluster.ord <- function(){
  order.county.clusters(mort.cluster.raw(), mort.cluster.map())
}


# Weighed Avg by ORDERED cluster
# Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  
  # Notes:
  #   - The cluster labels are ORDERED
mort.avg.cluster.ord <- function(){
  dplyr::mutate(mort.avg.cluster.raw(), cluster = mort.cluster.map()[cluster, "ord"])
}

#Extracting the mortality rates for all states and counties for years "2015-2017"
mort.rate <- function(){
  cdc.data %>% dplyr::filter(
    death_cause == death_cause_cardiovascular,
    period == "2015-2017"
  ) %>%
    dplyr::mutate(
      # death_rate = death_num / population * 10^5
      #death_rate = cut(death_rate, bin.geo.mort(death_cause_cardiovascular))
    ) %>%
    dplyr::select(county_fips, death_rate)
  
}
# Kendall Correlation Between raw mortality rates and county health rankings social determinants
kendall.cor <- mort.rate() %>% 
      dplyr::mutate(VAR = death_rate) %>%
      kendall.func(chr.data.2019) %>%
      dplyr::mutate(
        DIR = dplyr::if_else(
          kendall_cor <= 0,
          "Protective",
          "Destructive"
        )
        #chr_code = chr.namemap.2019[chr_code, 1]
      ) %>% na.omit() %>% 
      dplyr::arrange(desc(kendall_cor)) %>% 
      dplyr::mutate(chr_code = reorder(chr_code, kendall_cor))

selected_SDs_cardiovascular <- cbind(kendall.cor, state = state)

#using Benjamin Hochberg's p.adjust() method to adjust P-values for multiple hypothesis testing to filter out factors that are less relevant
corrected_pvalues <- p.adjust(selected_SDs_cardiovascular$kendall_p)
selected_SDs_cardiovascular <- cbind(selected_SDs_cardiovascular, Corrected_P_Vals = corrected_pvalues)
selected_SDs_cardiovascular <- selected_SDs_cardiovascular %>%
  dplyr::filter(Corrected_P_Vals < 0.05) %>% 
  dplyr::arrange(desc(kendall_cor))

#Printing out the few selected factors for cardiovascular deaths
print(selected_SDs_cardiovascular)
```


Filtering out social determinants that are less relevant for all causes of death using Benjamin Hochberg's method for multiple hypothesis testing.
```{r}
state = "US"
death_cause_allcause = "All Cause"

# UNORDERED mortality trend cluster label calculation
 # Variables:
  #   - county_fips
  #   - cluster
mort.cluster.raw <- function(){
  n.clusters <- 6
  cluster.counties(cdc.mort.mat(cdc.data, "US", death_cause_allcause),
                   cluster.method="kmeans",
                   cluster.num=n.clusters)
}


# Weighed Avg by UNORDERED cluster
# Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  
  # Notes:
  #   - The cluster labels are UNORDERED
mort.avg.cluster.raw <- function(){
  head(cdc.data)
  cdc.data %>%
    dplyr::filter(death_cause == death_cause_allcause) %>%
    dplyr::right_join(mort.cluster.raw(), by = "county_fips") %>%
    dplyr::group_by(period, cluster) %>%
    dplyr::summarise(
      death_rate = sum(death_num) / sum(population) * 10^5,
      count = n()
    ) %>% 
    dplyr::ungroup()
}


# MAPPING from UNORDERED mortality trend label to ORDERED mortality trend label
# Variables:
  #   - ord
  
  # Notes:
  #   - This is a mapping from raw cluster label to ORDERED cluster.
  #       Row names are the original cluster and `ord` are the reordered cluster
mort.cluster.map <- function(){
  mort.avg.cluster.raw() %>% 
    dplyr::filter(period == "2015-2017") %>%
    dplyr::arrange(death_rate) %>% 
    dplyr::mutate(ord = as.character(1:n())) %>% 
    dplyr::select(-c(period, death_rate)) %>% 
    textshape::column_to_rownames("cluster")
}


# ORDERED mortality trend cluster label calculation
 # Variables:
  #   - county_fips
  #   - cluster
mort.cluster.ord <- function(){
  order.county.clusters(mort.cluster.raw(), mort.cluster.map())
}


# Weighed Avg by ORDERED cluster
# Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  
  # Notes:
  #   - The cluster labels are ORDERED
mort.avg.cluster.ord <- function(){
  dplyr::mutate(mort.avg.cluster.raw(), cluster = mort.cluster.map()[cluster, "ord"])
}


mort.rate <- function(){
  
  cdc.data %>% dplyr::filter(
    death_cause == death_cause_allcause,
    period == "2015-2017"
  ) %>%
    dplyr::mutate(
      # death_rate = death_num / population * 10^5
      #death_rate = cut(death_rate, bin.geo.mort(death_cause_allcause))
    ) %>%
    dplyr::select(county_fips, death_rate)
  
}

# Kendall Correlation Between raw mortality rates and county health rankings social determinants
kendall.cor <- mort.rate() %>% 
      dplyr::mutate(VAR = death_rate) %>%
      kendall.func(chr.data.2019) %>%
      dplyr::mutate(
        DIR = dplyr::if_else(
          kendall_cor <= 0,
          "Protective",
          "Destructive"
        )
        #chr_code = chr.namemap.2019[chr_code, 1]
      ) %>% na.omit() %>% 
      dplyr::arrange(desc(kendall_cor)) %>% 
      dplyr::mutate(chr_code = reorder(chr_code, kendall_cor))

selected_SDs_allcause <- cbind(kendall.cor, state = state)

#using Benjamin Hochberg's p.adjust() method to adjust P-values for multiple hypothesis testing to filter out factors that are less relevant.
corrected_pvalues <- p.adjust(selected_SDs_allcause$kendall_p)
selected_SDs_allcause <- cbind(selected_SDs_allcause, Corrected_P_Vals = corrected_pvalues)
selected_SDs_allcause <- selected_SDs_allcause %>%
  dplyr::filter(Corrected_P_Vals < 0.05) %>% 
  dplyr::arrange(desc(kendall_cor))

#Printing out the few selected factors for all causes of death
print(selected_SDs_allcause)
```


Combining all the influential social determinants that are are found in all causes of death. We are combining them by taking the union.
```{r}
selected_SDs_final <- union(selected_SDs_despair$chr_code, selected_SDs_cancer$chr_code)
selected_SDs_final <- union(selected_SDs_final, selected_SDs_assault$chr_code)
selected_SDs_final <- union(selected_SDs_final, selected_SDs_cardiovascular$chr_code)
selected_SDs_final <- union(selected_SDs_final, selected_SDs_allcause$chr_code)
print(selected_SDs_final)
```

Creating a dataframe that only contains the final selected social determinants
```{r}
selected_SDs_final <- selected_SDs_final[-1]
saveRDS(selected_SDs_final, "init/reduced.chr.list")

reduced.chr.list <- readRDS("init/reduced.chr.list")
reduced.chr.list <- append(reduced.chr.list, "county_fips", after = 0)
chr.data.2019_fewer_SDs <- chr.data.2019 %>%
  as_data_frame %>%
  select(reduced.chr.list)
head(chr.data.2019_fewer_SDs)

saveRDS(chr.data.2019_fewer_SDs, "init/chr.data.2019.reduced.rds")
```

```{r}
reduced.chr.list <- readRDS("init/chr.data.2019.reduced.rds")
print(reduced.chr.list)
```

Calculating the percentage missing values for each factor (social determinant)
```{r}
colMeans(is.na(chr.data.2019))
```

