---
title: "Statewide Kendall's Taus Extractor"
author: "Lilian Ngweta"
date: "09/27/2019"
output: html_notebook
---
Note: Re-used code written by: Yuxuan Wang, Shengjin Li, and Ziyi Wang



Reading Source.R
```{r}
#source("Source.R")

chr.data.2019 <- readRDS("init/chr.data.2019.rds") %>%

  # 10/02/2019
  # Remove unwanted social determinants
  # This should be a temporary change because the source file will be replaced in the end
  dplyr::select(
    -dplyr::starts_with("health_"),
    -dplyr::contains("cohort"),
    -dplyr::contains("firearm"),
    -dplyr::contains("mortality"),
    -dplyr::contains("deaths"),
    -dplyr::contains("death"),
    -dplyr::contains("life"),
    -dplyr::contains("ypll"),
    -dplyr::starts_with("num_")
  )
```


Extracting Taus for all social determinants on nationwide data. The goal is to filter out social dterminants that are less relevant.

For Deaths of Despair
```{r}
state = "US"
death_cause_despair = "Despair"
# UNORDERED mortality trend cluster label calculation
mort.cluster.raw <- function(){
  
  # Variables:
  #   - county_fips
  #   - cluster
  
  # Currently hard-coded 4 clusters
  n.clusters <- 6
  cluster.counties(cdc.mort.mat(cdc.data, "US", death_cause_despair),
                   cluster.method="kmeans",
                   cluster.num=n.clusters)
}

# Weighed Avg by UNORDERED cluster
mort.avg.cluster.raw <- function(){
  
  # Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  
  # Notes:
  #   - The cluster labels are UNORDERED
  head(cdc.data)
  cdc.data %>%
    dplyr::filter(death_cause == death_cause_despair) %>%
    dplyr::right_join(mort.cluster.raw(), by = "county_fips") %>%
    dplyr::group_by(period, cluster) %>%
    dplyr::summarise(
      death_rate = sum(death_num) / sum(population) * 10^5,
      count = n()
    ) %>% 
    dplyr::ungroup()
}

# MAPPING from UNORDERED mortality trend label to ORDERED mortality trend label
mort.cluster.map <- function(){
  
  # Variables:
  #   - ord
  
  # Notes:
  #   - This is a mapping from raw cluster label to ORDERED cluster.
  #       Row names are the original cluster and `ord` are the reordered cluster
  
  mort.avg.cluster.raw() %>% 
    dplyr::filter(period == "2015-2017") %>%
    dplyr::arrange(death_rate) %>% 
    dplyr::mutate(ord = as.character(1:n())) %>% 
    dplyr::select(-c(period, death_rate)) %>% 
    textshape::column_to_rownames("cluster")
}

# ORDERED mortality trend cluster label calculation
mort.cluster.ord <- function(){
  
  # Variables:
  #   - county_fips
  #   - cluster
  
  order.county.clusters(mort.cluster.raw(), mort.cluster.map())
  #dplyr::mutate(mort.cluster.raw(), cluster = mort.cluster.map()[cluster, "ord"])
}

# Weighed Avg by ORDERED cluster
mort.avg.cluster.ord <- function(){
  
  # Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  
  # Notes:
  #   - The cluster labels are ORDERED
  
  dplyr::mutate(mort.avg.cluster.raw(), cluster = mort.cluster.map()[cluster, "ord"])
}

mort.rate <- function(){
  
  cdc.data %>% dplyr::filter(
    death_cause == death_cause_despair,
    #state_abbr == input$state_choice,
    period == "2015-2017"
  ) %>%
    dplyr::mutate(
      # death_rate = death_num / population * 10^5
      #death_rate = cut(death_rate, bin.geo.mort(death_cause_despair))
    ) %>%
    dplyr::select(county_fips, death_rate)
  
}


#print(mort.cluster.ord())
# Kendall Correlation Between raw mortality rates and county health rankings social determinants

#kendall.cor <- kendall.func(mort.cluster.ord(), chr.data.2019)
#kendall.cor <- kendall.func(mort.rate(), chr.data.2019)


kendall.cor <- mort.rate() %>% 
      dplyr::mutate(VAR = death_rate) %>%
      kendall.func(chr.data.2019) %>%
      dplyr::mutate(
        DIR = dplyr::if_else(
          kendall_cor <= 0,
          "Protective",
          "Destructive"
        )
        #chr_code = chr.namemap.2019[chr_code, 1]
      ) %>% na.omit() %>% 
      #dplyr::filter(kendall_p < 0.1) %>% 
      dplyr::arrange(desc(kendall_cor)) %>% 
      #dplyr::top_n(15, kendall_cor) %>% 
      dplyr::mutate(chr_code = reorder(chr_code, kendall_cor))
    



# kendall.cor <- kendall.func(chr.data.2019)
# 
# kendall.cor <- kendall.cor %>%
#   dplyr::mutate(
#     Direction = dplyr::if_else(
#       kendall_cor <= 0,
#       "Protective",
#       "Destructive"
#     ),
#     #kendall_cor = abs(kendall_cor),
#     kendall_cor = kendall_cor
#     #chr_code = chr.namemap.2019[chr_code, 1]
#   ) %>% na.omit() %>% 
#   #kendall_p = p.adjust(kendall_p, method = p.adjust.methods, n = length(kendall_p)) %>% 
#   dplyr::filter(kendall_p < 0.05) %>% 
#   #dplyr::filter(p.adjust(kendall_p, method = p.adjust.methods, n = length(kendall_p))) %>% 
#   dplyr::arrange(desc(kendall_cor))

selected_SDs_despair <- cbind(kendall.cor, state = state)

#using Benjamin Hochberg's p.adjust() method to adjust P-values for multiple hypothesis testing
corrected_pvalues <- p.adjust(selected_SDs_despair$kendall_p)

selected_SDs_despair <- cbind(selected_SDs_despair, Corrected_P_Vals = corrected_pvalues)

selected_SDs_despair <- selected_SDs_despair %>%
  #kendall_p = p.adjust(kendall_p, method = p.adjust.methods, n = length(kendall_p)) %>% 
  dplyr::filter(Corrected_P_Vals < 0.05) %>% 
  #dplyr::filter(p.adjust(kendall_p, method = p.adjust.methods, n = length(kendall_p))) %>% 
  dplyr::arrange(desc(kendall_cor))

# selected_SDs2[selected_SDs2$kendall_p == p.adjust(selected_SDs2$kendall_p),]
print(selected_SDs_despair)
```


For Cancer
```{r}
state = "US"

death_cause_cancer = "Cancer"
# UNORDERED mortality trend cluster label calculation
mort.cluster.raw <- function(){
  
  # Variables:
  #   - county_fips
  #   - cluster
  
  # Currently hard-coded 4 clusters
  n.clusters <- 6
  cluster.counties(cdc.mort.mat(cdc.data, "US", death_cause_cancer),
                   cluster.method="kmeans",
                   cluster.num=n.clusters)
}

# Weighed Avg by UNORDERED cluster
mort.avg.cluster.raw <- function(){
  
  # Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  
  # Notes:
  #   - The cluster labels are UNORDERED
  head(cdc.data)
  cdc.data %>%
    dplyr::filter(death_cause == death_cause_cancer) %>%
    dplyr::right_join(mort.cluster.raw(), by = "county_fips") %>%
    dplyr::group_by(period, cluster) %>%
    dplyr::summarise(
      death_rate = sum(death_num) / sum(population) * 10^5,
      count = n()
    ) %>% 
    dplyr::ungroup()
}

# MAPPING from UNORDERED mortality trend label to ORDERED mortality trend label
mort.cluster.map <- function(){
  
  # Variables:
  #   - ord
  
  # Notes:
  #   - This is a mapping from raw cluster label to ORDERED cluster.
  #       Row names are the original cluster and `ord` are the reordered cluster
  
  mort.avg.cluster.raw() %>% 
    dplyr::filter(period == "2015-2017") %>%
    dplyr::arrange(death_rate) %>% 
    dplyr::mutate(ord = as.character(1:n())) %>% 
    dplyr::select(-c(period, death_rate)) %>% 
    textshape::column_to_rownames("cluster")
}

# ORDERED mortality trend cluster label calculation
mort.cluster.ord <- function(){
  
  # Variables:
  #   - county_fips
  #   - cluster
  
  order.county.clusters(mort.cluster.raw(), mort.cluster.map())
  #dplyr::mutate(mort.cluster.raw(), cluster = mort.cluster.map()[cluster, "ord"])
}

# Weighed Avg by ORDERED cluster
mort.avg.cluster.ord <- function(){
  
  # Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  
  # Notes:
  #   - The cluster labels are ORDERED
  
  dplyr::mutate(mort.avg.cluster.raw(), cluster = mort.cluster.map()[cluster, "ord"])
}

mort.rate <- function(){
  
  cdc.data %>% dplyr::filter(
    death_cause == death_cause_cancer,
    #state_abbr == input$state_choice,
    period == "2015-2017"
  ) %>%
    dplyr::mutate(
      # death_rate = death_num / population * 10^5
      #death_rate = cut(death_rate, bin.geo.mort(death_cause_cancer))
    ) %>%
    dplyr::select(county_fips, death_rate)
  
}


#print(mort.cluster.ord())
# Kendall Correlation Between raw mortality rates and county health rankings social determinants

#kendall.cor <- kendall.func(mort.cluster.ord(), chr.data.2019)
# kendall.cor <- kendall.func(mort.rate(), chr.data.2019)
# 
# kendall.cor <- kendall.cor %>%
#   dplyr::mutate(
#     Direction = dplyr::if_else(
#       kendall_cor <= 0,
#       "Protective",
#       "Destructive"
#     ),
#     #kendall_cor = abs(kendall_cor),
#     kendall_cor = kendall_cor
#     #chr_code = chr.namemap.2019[chr_code, 1]
#   ) %>% na.omit() %>% 
#   #kendall_p = p.adjust(kendall_p, method = p.adjust.methods, n = length(kendall_p)) %>% 
#   dplyr::filter(kendall_p < 0.05) %>% 
#   #dplyr::filter(p.adjust(kendall_p, method = p.adjust.methods, n = length(kendall_p))) %>% 
#   dplyr::arrange(desc(kendall_cor))


kendall.cor <- mort.rate() %>% 
      dplyr::mutate(VAR = death_rate) %>%
      kendall.func(chr.data.2019) %>%
      dplyr::mutate(
        DIR = dplyr::if_else(
          kendall_cor <= 0,
          "Protective",
          "Destructive"
        )
        #chr_code = chr.namemap.2019[chr_code, 1]
      ) %>% na.omit() %>% 
      #dplyr::filter(kendall_p < 0.1) %>% 
      dplyr::arrange(desc(kendall_cor)) %>% 
      #dplyr::top_n(15, kendall_cor) %>% 
      dplyr::mutate(chr_code = reorder(chr_code, kendall_cor))


selected_SDs_cancer <- cbind(kendall.cor, state = state)

#using Benjamin Hochberg's p.adjust() method to adjust P-values for multiple hypothesis testing
corrected_pvalues <- p.adjust(selected_SDs_cancer$kendall_p)

selected_SDs_cancer <- cbind(selected_SDs_cancer, Corrected_P_Vals = corrected_pvalues)

selected_SDs_cancer <- selected_SDs_cancer %>%
  #kendall_p = p.adjust(kendall_p, method = p.adjust.methods, n = length(kendall_p)) %>% 
  dplyr::filter(Corrected_P_Vals < 0.05) %>% 
  #dplyr::filter(p.adjust(kendall_p, method = p.adjust.methods, n = length(kendall_p))) %>% 
  dplyr::arrange(desc(kendall_cor))

# selected_SDs2[selected_SDs2$kendall_p == p.adjust(selected_SDs2$kendall_p),]
print(selected_SDs_cancer)
```



For Assault
```{r}
state = "US"

death_cause_assault = "Assault"
# UNORDERED mortality trend cluster label calculation
mort.cluster.raw <- function(){
  
  # Variables:
  #   - county_fips
  #   - cluster
  
  # Currently hard-coded 4 clusters
  n.clusters <- 6
  cluster.counties(cdc.mort.mat(cdc.data, "US", death_cause_assault),
                   cluster.method="kmeans",
                   cluster.num=n.clusters)
}

# Weighed Avg by UNORDERED cluster
mort.avg.cluster.raw <- function(){
  
  # Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  
  # Notes:
  #   - The cluster labels are UNORDERED
  head(cdc.data)
  cdc.data %>%
    dplyr::filter(death_cause == death_cause_assault) %>%
    dplyr::right_join(mort.cluster.raw(), by = "county_fips") %>%
    dplyr::group_by(period, cluster) %>%
    dplyr::summarise(
      death_rate = sum(death_num) / sum(population) * 10^5,
      count = n()
    ) %>% 
    dplyr::ungroup()
}

# MAPPING from UNORDERED mortality trend label to ORDERED mortality trend label
mort.cluster.map <- function(){
  
  # Variables:
  #   - ord
  
  # Notes:
  #   - This is a mapping from raw cluster label to ORDERED cluster.
  #       Row names are the original cluster and `ord` are the reordered cluster
  
  mort.avg.cluster.raw() %>% 
    dplyr::filter(period == "2015-2017") %>%
    dplyr::arrange(death_rate) %>% 
    dplyr::mutate(ord = as.character(1:n())) %>% 
    dplyr::select(-c(period, death_rate)) %>% 
    textshape::column_to_rownames("cluster")
}

# ORDERED mortality trend cluster label calculation
mort.cluster.ord <- function(){
  
  # Variables:
  #   - county_fips
  #   - cluster
  
  order.county.clusters(mort.cluster.raw(), mort.cluster.map())
  #dplyr::mutate(mort.cluster.raw(), cluster = mort.cluster.map()[cluster, "ord"])
}

# Weighed Avg by ORDERED cluster
mort.avg.cluster.ord <- function(){
  
  # Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  
  # Notes:
  #   - The cluster labels are ORDERED
  
  dplyr::mutate(mort.avg.cluster.raw(), cluster = mort.cluster.map()[cluster, "ord"])
}

mort.rate <- function(){
  
  cdc.data %>% dplyr::filter(
    death_cause == death_cause_assault,
    #state_abbr == input$state_choice,
    period == "2015-2017"
  ) %>%
    dplyr::mutate(
      # death_rate = death_num / population * 10^5
      #death_rate = cut(death_rate, bin.geo.mort(death_cause_assault))
    ) %>%
    dplyr::select(county_fips, death_rate)
  
}


#print(mort.cluster.ord())
# Kendall Correlation Between raw mortality rates and county health rankings social determinants

#kendall.cor <- kendall.func(mort.cluster.ord(), chr.data.2019)
# kendall.cor <- kendall.func(mort.rate(), chr.data.2019)
# 
# kendall.cor <- kendall.cor %>%
#   dplyr::mutate(
#     Direction = dplyr::if_else(
#       kendall_cor <= 0,
#       "Protective",
#       "Destructive"
#     ),
#     #kendall_cor = abs(kendall_cor),
#     kendall_cor = kendall_cor
#     #chr_code = chr.namemap.2019[chr_code, 1]
#   ) %>% na.omit() %>% 
#   #kendall_p = p.adjust(kendall_p, method = p.adjust.methods, n = length(kendall_p)) %>% 
#   dplyr::filter(kendall_p < 0.05) %>% 
#   #dplyr::filter(p.adjust(kendall_p, method = p.adjust.methods, n = length(kendall_p))) %>% 
#   dplyr::arrange(desc(kendall_cor))

kendall.cor <- mort.rate() %>% 
      dplyr::mutate(VAR = death_rate) %>%
      kendall.func(chr.data.2019) %>%
      dplyr::mutate(
        DIR = dplyr::if_else(
          kendall_cor <= 0,
          "Protective",
          "Destructive"
        )
        #chr_code = chr.namemap.2019[chr_code, 1]
      ) %>% na.omit() %>% 
      #dplyr::filter(kendall_p < 0.1) %>% 
      dplyr::arrange(desc(kendall_cor)) %>% 
      #dplyr::top_n(15, kendall_cor) %>% 
      dplyr::mutate(chr_code = reorder(chr_code, kendall_cor))

selected_SDs_assault <- cbind(kendall.cor, state = state)

#using Benjamin Hochberg's p.adjust() method to adjust P-values for multiple hypothesis testing
corrected_pvalues <- p.adjust(selected_SDs_assault$kendall_p)

selected_SDs_assault <- cbind(selected_SDs_assault, Corrected_P_Vals = corrected_pvalues)

selected_SDs_assault <- selected_SDs_assault %>%
  #kendall_p = p.adjust(kendall_p, method = p.adjust.methods, n = length(kendall_p)) %>% 
  dplyr::filter(Corrected_P_Vals < 0.05) %>% 
  #dplyr::filter(p.adjust(kendall_p, method = p.adjust.methods, n = length(kendall_p))) %>% 
  dplyr::arrange(desc(kendall_cor))

# selected_SDs2[selected_SDs2$kendall_p == p.adjust(selected_SDs2$kendall_p),]
print(selected_SDs_assault)
```



For Cardiovascular
```{r}
state = "US"

death_cause_cardiovascular = "Cardiovascular"
# UNORDERED mortality trend cluster label calculation
mort.cluster.raw <- function(){
  
  # Variables:
  #   - county_fips
  #   - cluster
  
  # Currently hard-coded 4 clusters
  n.clusters <- 6
  cluster.counties(cdc.mort.mat(cdc.data, "US", death_cause_cardiovascular),
                   cluster.method="kmeans",
                   cluster.num=n.clusters)
}

# Weighed Avg by UNORDERED cluster
mort.avg.cluster.raw <- function(){
  
  # Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  
  # Notes:
  #   - The cluster labels are UNORDERED
  head(cdc.data)
  cdc.data %>%
    dplyr::filter(death_cause == death_cause_cardiovascular) %>%
    dplyr::right_join(mort.cluster.raw(), by = "county_fips") %>%
    dplyr::group_by(period, cluster) %>%
    dplyr::summarise(
      death_rate = sum(death_num) / sum(population) * 10^5,
      count = n()
    ) %>% 
    dplyr::ungroup()
}

# MAPPING from UNORDERED mortality trend label to ORDERED mortality trend label
mort.cluster.map <- function(){
  
  # Variables:
  #   - ord
  
  # Notes:
  #   - This is a mapping from raw cluster label to ORDERED cluster.
  #       Row names are the original cluster and `ord` are the reordered cluster
  
  mort.avg.cluster.raw() %>% 
    dplyr::filter(period == "2015-2017") %>%
    dplyr::arrange(death_rate) %>% 
    dplyr::mutate(ord = as.character(1:n())) %>% 
    dplyr::select(-c(period, death_rate)) %>% 
    textshape::column_to_rownames("cluster")
}

# ORDERED mortality trend cluster label calculation
mort.cluster.ord <- function(){
  
  # Variables:
  #   - county_fips
  #   - cluster
  
  order.county.clusters(mort.cluster.raw(), mort.cluster.map())
  #dplyr::mutate(mort.cluster.raw(), cluster = mort.cluster.map()[cluster, "ord"])
}

# Weighed Avg by ORDERED cluster
mort.avg.cluster.ord <- function(){
  
  # Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  
  # Notes:
  #   - The cluster labels are ORDERED
  
  dplyr::mutate(mort.avg.cluster.raw(), cluster = mort.cluster.map()[cluster, "ord"])
}

mort.rate <- function(){
  
  cdc.data %>% dplyr::filter(
    death_cause == death_cause_cardiovascular,
    #state_abbr == input$state_choice,
    period == "2015-2017"
  ) %>%
    dplyr::mutate(
      # death_rate = death_num / population * 10^5
      #death_rate = cut(death_rate, bin.geo.mort(death_cause_cardiovascular))
    ) %>%
    dplyr::select(county_fips, death_rate)
  
}


#print(mort.cluster.ord())
# Kendall Correlation Between raw mortality rates and county health rankings social determinants

#kendall.cor <- kendall.func(mort.cluster.ord(), chr.data.2019)
# kendall.cor <- kendall.func(mort.rate(), chr.data.2019)
# 
# kendall.cor <- kendall.cor %>%
#   dplyr::mutate(
#     Direction = dplyr::if_else(
#       kendall_cor <= 0,
#       "Protective",
#       "Destructive"
#     ),
#     #kendall_cor = abs(kendall_cor),
#     kendall_cor = kendall_cor
#     #chr_code = chr.namemap.2019[chr_code, 1]
#   ) %>% na.omit() %>% 
#   #kendall_p = p.adjust(kendall_p, method = p.adjust.methods, n = length(kendall_p)) %>% 
#   dplyr::filter(kendall_p < 0.05) %>% 
#   #dplyr::filter(p.adjust(kendall_p, method = p.adjust.methods, n = length(kendall_p))) %>% 
#   dplyr::arrange(desc(kendall_cor))

kendall.cor <- mort.rate() %>% 
      dplyr::mutate(VAR = death_rate) %>%
      kendall.func(chr.data.2019) %>%
      dplyr::mutate(
        DIR = dplyr::if_else(
          kendall_cor <= 0,
          "Protective",
          "Destructive"
        )
        #chr_code = chr.namemap.2019[chr_code, 1]
      ) %>% na.omit() %>% 
      #dplyr::filter(kendall_p < 0.1) %>% 
      dplyr::arrange(desc(kendall_cor)) %>% 
      #dplyr::top_n(15, kendall_cor) %>% 
      dplyr::mutate(chr_code = reorder(chr_code, kendall_cor))

selected_SDs_cardiovascular <- cbind(kendall.cor, state = state)

#using Benjamin Hochberg's p.adjust() method to adjust P-values for multiple hypothesis testing
corrected_pvalues <- p.adjust(selected_SDs_cardiovascular$kendall_p)

selected_SDs_cardiovascular <- cbind(selected_SDs_cardiovascular, Corrected_P_Vals = corrected_pvalues)

selected_SDs_cardiovascular <- selected_SDs_cardiovascular %>%
  #kendall_p = p.adjust(kendall_p, method = p.adjust.methods, n = length(kendall_p)) %>% 
  dplyr::filter(Corrected_P_Vals < 0.05) %>% 
  #dplyr::filter(p.adjust(kendall_p, method = p.adjust.methods, n = length(kendall_p))) %>% 
  dplyr::arrange(desc(kendall_cor))

# selected_SDs2[selected_SDs2$kendall_p == p.adjust(selected_SDs2$kendall_p),]
print(selected_SDs_cardiovascular)
```



Influential social determinants that are are found in all causes of death
```{r}
selected_SDs <- intersect(selected_SDs_despair$chr_code, selected_SDs_cancer$chr_code)
selected_SDs <- intersect(selected_SDs, selected_SDs_assault$chr_code)
selected_SDs <- intersect(selected_SDs, selected_SDs_cardiovascular$chr_code)
print(selected_SDs)
```


Combining all the influential social determinants that are are found in all causes of death
```{r}
selected_SDs_final <- union(selected_SDs_despair$chr_code, selected_SDs_cancer$chr_code)
selected_SDs_final <- union(selected_SDs_final, selected_SDs_assault$chr_code)
selected_SDs_final <- union(selected_SDs_final, selected_SDs_cardiovascular$chr_code)
print(selected_SDs_final)
```

Creating a dataframe that only contains the final selected social determinants
```{r}
# chr.data.2019 %>% 
#   as_data_frame %>%
#   select_(.dots = selected_SDs_final)

selected_SDs_final <- selected_SDs_final[-1]
selected_SDs_final <- append(selected_SDs_final, "county_fips", after = 0)

chr.data.2019_fewer_SDs <- chr.data.2019 %>%
  as_data_frame %>%
  select(selected_SDs_final)


#chr.data.2019[,selected_SDs_final]
head(chr.data.2019_fewer_SDs)

saveRDS(chr.data.2019_fewer_SDs, "init/chr.data.2019.reduced.rds")
```


```{r}
head(chr.data.2019)
```

