---
title: "Run App Analysis"
author: "Ross DeVito"
date: "October 1, 2019"
output:
  pdf_document: default
  html_document: default
  pdf_output: default
toc: default
---

# Overview

This notebook loads in data sources and functions used in app and applies to selected state and death cause so that all data used in the app at the time those variable are selected is available here.

# Libraries and Environment
```{r setup, include=FALSE}
set.seed(300)
current.dir <- getwd()

# set working directory to current apps init
path.to.current.app <- file.path("..", "init")
setwd(file.path(path.to.current.app))

source("Librarian.R")
source("Loader_CHR2019.R")
source("Loader_CDC.R")
source("Loader_GEO.R")
source("Analyzer_PCA.R")
source("Clustering_Lib.R")
source("Analyzer_Correlation.R")
source("Theme.R") # for styles for plots needed in run_app_analysis, could be optional

# revert to original working directory
#setwd(current.dir)
```

# Function definitions

## Cache of UNORDERED mortality trend cluster label calculation.

* Currently hard-coded 4 clusters
```{r}
mort.cluster.raw <- function() {
  # Variables:
  #   - county_fips
  #   - cluster
  
  cdc.data %>%
    cdc.mort.mat(input$state_choice, input$death_cause) %>%
    km.func(4)
}
```

## Cache of Weighed Avg by UNORDERED cluster

* The cluster labels are UNORDERED
```{r}
mort.avg.cluster.raw <- function() {

  # Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count
  cdc.data %>%
    dplyr::filter(state_abbr == input$state_choice, death_cause == input$death_cause) %>%
    dplyr::right_join(mort.cluster.raw(), by = "county_fips") %>%
    dplyr::group_by(period, cluster) %>%
    dplyr::summarise(
      death_rate = sum(death_num) / sum(population) * 10^5,
      count = n()
    ) %>%
    dplyr::ungroup()
}
```

## Cache of MAPPING from UNORDERED mortality trend label to ORDERED mortality trend label

* This is a mapping from raw cluster label to ORDERED cluster.
* Row names are the original cluster and `ord` are the reordered cluster
```{r}
mort.cluster.map <- function() {
  # Variables:
  #   - ord
  mort.avg.cluster.raw() %>%
    dplyr::filter(period == "2015-2017") %>%
    dplyr::arrange(death_rate) %>%
    dplyr::mutate(ord = as.character(1:n())) %>%
    dplyr::select(-c(period, death_rate)) %>%
    textshape::column_to_rownames("cluster")
}
```

## Cache of ORDERED mortality trend cluster label calculation

```{r}
mort.cluster.ord <- function() {
  # Variables:
  #   - county_fips
  #   - cluster
  dplyr::mutate(mort.cluster.raw(), cluster = mort.cluster.map()[cluster, "ord"])
}
```

## Cache of Weighed Avg by ORDERED cluster

* The cluster labels are ORDERED
```{r}
mort.avg.cluster.ord <- function() {
  # Variables:
  #   - period
  #   - cluster
  #   - death_rate
  #   - count

  dplyr::mutate(mort.avg.cluster.raw(), cluster = mort.cluster.map()[cluster, "ord"])
}
```

# Parameters

```{r}
death.causes <- c("Despair","Cancer","Assault","Cardiovascular")
input <- list(state_choice = state.abb[1],
              death_cause = death.causes[1])
```

# Mortality Rate Trend Line Graph

```{r}
plot_mort_line <- function() {
  ggplot(
    mort.avg.cluster.ord(),
    aes(
      x = period, y = death_rate,
      color = cluster, group = cluster
    )
  ) +
    geom_line(size = 1) +
    geom_point(color = "black", shape = 21, fill = "white") +
    labs.line.mort(input$state_choice, input$death_cause) +
    color.line.cluster() +
    theme.line.mort() +
    guides(
      color = guide_legend(reverse = T)
    )
}
#plot_mort_line()  <- Not sure how to run this
```

# Mortality Rate Table

```{r}
get.cluster.table <- function() {
  rate.table <- mort.avg.cluster.ord() %>%
    dplyr::select(cluster, period, death_rate) %>%
    tidyr::spread(key = period, value = death_rate) %>%
    dplyr::select(cluster, `2000-2002`, `2015-2017`)

  count.table <- mort.avg.cluster.ord() %>%
    dplyr::select(cluster, count) %>%
    base::unique()

  dplyr::left_join(count.table, rate.table, by = "cluster") %>%
    dplyr::mutate(cluster = as.character(cluster)) %>%
    dplyr::arrange(desc(cluster)) %>%
    dplyr::rename(
      "Trend Grp." = "cluster",
      "Count" = "count"
    )
}
cluster.table <- get.cluster.table()
knitr::kable(cluster.table)
```

```{r}
kendall.cor <- kendall.func(mort.cluster.ord(), chr.data.2019)
knitr::kable(kendall.cor)
```

