---
title: "social_determinant_zoom_in"
output: html_notebook
---

Shows data flow to zoom in on a social determinant for a state and display box plots or whatever other plot is decided on.

Load data and load libraries
```{r}
# save current directory so can return to it after running Loaders/Analyzers
starting.dir <- getwd()

# set working directory to current app
path.to.current.app <- file.path("..")
setwd(path.to.current.app)

source("Source.R")

# notebook automatically returns to starting.dir
```

Select state and cause of death, then get correlations

```{r}
state <- "AL"
death.cause <- "Despair"

county.clusters <- cluster.counties(cdc.mort.mat(cdc.data, state, death.cause),
                                    cluster.method="kmeans",
                                    cluster.num=3)
cluster.deathrates <- get.cluster.deathrate.during.time(county.clusters, cdc.data, death.cause)
cluster.order.map <- get.cluster.order.map(cluster.deathrates)
county.clusters.ordered <- order.county.clusters(county.clusters, cluster.order.map)
cluster.deathrates.ordered <- order.cluster.deathrate.during.time(cluster.deathrates, cluster.order.map)

# get correlations
kendall.cor <- kendall.func(county.clusters.ordered, chr.data.2019)

# filter out determinants with p value above .05
kendall.cor <- kendall.cor[kendall.cor$kendall_p <= .05,]
```

Select a social determinant. In app will be done by clicking a dterminant bar

```{r}
selected.determinant <- kendall.cor$chr_code[4]
print(selected.determinant)
```

Create box plots for social determinants wrt county clusters. Note consistent colors between here and scatter plot

```{r}
cluster.determinant.vals <- dplyr::inner_join(county.clusters.ordered, 
                                              chr.data.2019[c("county_fips", selected.determinant)],
                                              by = "county_fips")

cluster.boxplots <- ggplot(cluster.determinant.vals)  + 
  geom_boxplot(aes_string(x = "cluster", y = selected.determinant, color = "cluster")) +
  xlab("Cluster") +
  ylab(chr.namemap.2019[selected.determinant,]) +
  theme(legend.position = "none")

cluster.boxplots
```

Create scatter plot of determinant vs morality with points colored by cluster

```{r}
# for scatter need to append most recent deathrate
cluster.determinant.vals.deathrate <- dplyr::inner_join(cluster.determinant.vals, 
                                              cdc.mort.mat(cdc.data, state, death.cause
                                                            )[c("county_fips", "2015-2017")],
                                              by = "county_fips")

colnames(cluster.determinant.vals.deathrate)[colnames(cluster.determinant.vals.deathrate)=="2015-2017"] <- "mortality_rate"

cluster.scatter <- ggplot(cluster.determinant.vals.deathrate)  + 
  geom_point(aes_string(x = selected.determinant, y = "mortality_rate", color = "cluster")) +
  xlab(chr.namemap.2019[selected.determinant,]) +
  ylab("Mortality Rate") +
  labs(color = "Cluster\n")

cluster.scatter
```


