---
title: "State Comparison"
Author: Farukh Saidmuratov 
Date: Oct 16, 2019 
---

# Libraries and Environment
```{r}
set.seed(300)
current.dir <- getwd()

# set working directory to current apps init
path.to.current.app <- file.path("..", "init")
setwd(file.path(path.to.current.app))

source("Librarian.R")
source("Loader_CHR2019.R")
source("Loader_CDC.R")
source("Loader_GEO.R")
source("Analyzer_PCA.R")
source("Clustering_Lib.R")
source("Analyzer_Correlation.R")
source("Theme.R") # for styles for plots 

source("CDC_Lib.R")# For making plot 

# revert to original working directory
current.dir <- getwd()
setwd(current.dir)
```

# Labs line mort for multiple states 
```{r}
labs.line.mort.mult <- function(state.choice1, state.choice2, death.cause) {
  labs(
    title = paste(
      state.choice1, 
      paste("Death of", death.cause, "Trend"), 
      state.choice2, 
      sep = ' - '
    ),
    caption = "Data Sorce: CDCWONDER Multi-Cause of Death",
    x = "Period", y = "Mortality Rate (# per 100k)"
  )
}
```

# Parameters. Set first state to be Mass., and second state to be Connecticut Explore difference between deaths of cardiovascular disease.
```{r}
death.causes <- c("Despair","Cancer","Assault","Cardiovascular")
input <- list(state_choice1 = state.abb[21], state_choice2 = state.abb[7], 
              death_cause = death.causes[4])
```

# Compare the mortality trends of Mass. and CT 2 New England states, where MA has Affordable Health Care
# Act type policies, while CT doesn't 

# Get data on states 
```{r}
data.state1 <- cdc.mort.mat(cdc.data, input$state_choice1, input$death_cause)
data.state2 <- cdc.mort.mat(cdc.data, input$state_choice2, input$death_cause)
```

# Sum mortality rate on time period (columns) and put into dataframe 
```{r}
periods <- c("2000-2002", "2003-2005", "2006-2008", "2009-2011", "2012-2014", "2015-2017")
state.compare <- data.frame("period" = character(), "state" = character(), "death_rate"=integer(), stringsAsFactors = FALSE)

for(period in periods) {
  state.compare[nrow(state.compare) + 1,] <- c(period, input$state_choice1, round(sum(data.state1[period]), 2))
  state.compare[nrow(state.compare) + 1,] <- c(period, input$state_choice2, round(sum(data.state2[period]), 2))
}
```

# Consider 2 states as 2 different "clusters"
```{r}
color.line.cluster <- function(state.choice, n.clusters) {
  
  if (state.choice != "US"){
    scale_color_manual(
      name = "Cluster",
      #c("#ffc4c4", "#ff8f8f", "#ff5454", "#ff1414", "#a80000")
      values = colorRampPalette(
        c("#feedde", "#fdd0a2", "#fdae6b", "#fd8d3c", "#e6550d", "#a63603")
        
      )(n.clusters),
      guide = guide_legend(reverse = T)
    )
    
  } else {
    scale_color_manual(
      name = "State",
      #c("#ffc4c4", "#ff8f8f", "#ff5454", "#ff1414", "#a80000")
      values = colorRampPalette(
        c("#fef0d9","#fdcc8a","#fc8d59","#e34a33")
        
      )(n.clusters),
      guide = guide_legend(reverse = T)
    )
  }
  
}
```


# Function to plot comparison between states. Change font family because of bug on my machine. 
```{r}
library(extrafont)
loadfonts(quiet = T)

plot_mort_line <- function() {
  ggplot(state.compare, aes(x = period, y = death_rate,color = state, group = state)) + geom_line(size = 1) + geom_point(color = "black", shape = 21, fill = "white") + labs.line.mort.mult(input$state_choice1, input$state_choice2, input$death_cause) + color.line.cluster(input$state_choice1, 2) + theme.line.mort() + guides(color = guide_legend(reverse = T)) + theme(text=element_text(family="Arial Black")) + scale_y_reverse(name="death_rate")
}
```

# Plot!
```{r}
plot_mort_line()
```


